Evolution of solar wind discontinuities in the inner heliosphere

### 2025-12-01 index.tex
Built site for gh-pages

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Evolution of solar wind discontinuities in the inner heliosphere},
  pdfauthor={Zijin Zhang; Anton V. Artemyev; Vassilis Angelopoulos; Tai Phan},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Evolution of solar wind discontinuities in the inner heliosphere}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{PSP and Earth conjunctions and alignments}
\author{Zijin Zhang \and Anton V. Artemyev \and Vassilis Angelopoulos \and Tai Phan}
\date{}
\begin{document}
\maketitle

\section{Abstract}\label{abstract}

We analyze three time intervals during which the Parker Solar Probe (PSP) was in close proximity to the Sun and investigate current sheets (magnetic discontinuities) observed by PSP, ARTEMIS, and Wind. Using an automated detection algorithm, we identify and examine the statistical properties of over 45,000 current sheets across the three intervals, including approximately 17,000 from PSP. The results show that current sheets broaden with increasing heliocentric distance, while their characteristic current density remains largely unchanged. The thickness and current density are strongly anti-correlated, indicating that thinner current sheets carry stronger currents, consistent with turbulent intermittency. Distinct distributions of magnetic field jumps and rotation angles are also observed among the different missions, suggesting structural evolution of current sheets as the solar wind expands. Most current sheets display small variations in magnetic field magnitude, indicating a predominantly force-free character. Significant deviations from the Walén relation persist even when pressure anisotropy is included, implying that anisotropy alone cannot explain the observed departures from ideal rotational discontinuities. Finally, the Alfvénicity of the current sheets correlates strongly with the cross helicity and residual energy of the surrounding solar wind, demonstrating that current sheet dynamics are largely governed by the ambient turbulent environment.

@@

Figure~\ref{fig-overview} provides an overview of the selected time interval from Encounter 7, comparing measurements from the PSP (left panels) and Wind (right panels) missions. From top to bottom, the panels display 8-minute averages of the magnetic field magnitude, proton number density, and proton bulk flow speed, followed by ion and electron temperatures resolved in the parallel and perpendicular directions. To facilitate comparison, Wind data are scaled to \(20 R_\odot\) using the empirical relations \(B_{\text{scaled}}/B = ( 20 R_\odot / 1\ \text{AU} )^{-1.59}\), \(n_{\text{scaled}}/n = ( 20 R_\odot / 1\ \text{AU} )^{-1.96}\) \citep{perroneRadialEvolutionSolar2019}.
The fifth and sixth panels show the median Alfvén ratio, \(R_{VB} = |Δ 𝐕| / |Δ 𝐕_A|\), of the current sheets and the total number of current sheets identified within each four-hour window. Including the anisotropy factor \(Λ = μ(p_{\parallel,e} − p_{\perp,e} + p_{\parallel,i} − p_{\perp,i})/B^2\) does not significantly affect the Alfvén ratio \(R_{VB}^* = R_{VB} / \sqrt{1 - Λ}\); its average value is approximately 0.5 for PSP and 0.4 for Wind over the entire time interval. The number of current sheets varies from 100 to 300 per four-hour window, corresponding to an average of about 1,000 and 900 current sheets per day for PSP and Wind, respectively.
Panel (a.7) and (b.7) present the cross helicity and residual energy, which describe the Alfvénic nature and energetic balance of the ambient plasma, respectively, calculated every 8 minutes. The final two panels display 8-minute averages of helium abundance and plasma beta, providing complementary information on the compositional and thermal conditions of the solar wind. These parameters offer critical insights into identifying the source regions of different solar wind streams \citep{altermanCrossHelicityHelium2025}.

>>>>

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Evolution of solar wind discontinuities in the inner heliosphere},
  pdfauthor={Zijin Zhang; Anton V. Artemyev; Vassilis Angelopoulos; Tai Phan},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Evolution of solar wind discontinuities in the inner heliosphere}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{PSP and Earth conjunctions and alignments}
\author{Zijin Zhang \and Anton V. Artemyev \and Vassilis Angelopoulos \and Tai Phan}
\date{}
\begin{document}
\maketitle
\begin{abstract}
We analyze three time intervals during which the Parker Solar Probe (PSP) was in close proximity to the Sun and investigate current sheets (magnetic discontinuities) observed by PSP, ARTEMIS, and Wind. Using an automated detection algorithm, we identify and examine the statistical properties of over 45,000 current sheets across the three intervals, including approximately 17,000 from PSP. The results show that current sheets broaden with increasing heliocentric distance, while their characteristic current density remains largely unchanged. The thickness and current density are strongly anti-correlated, indicating that thinner current sheets carry stronger currents, consistent with turbulent intermittency. Distinct distributions of magnetic field jumps and rotation angles are also observed among the different missions, suggesting structural evolution of current sheets as the solar wind expands. Most current sheets display small variations in magnetic field magnitude, indicating a predominantly force-free character. Significant deviations from the Walén relation persist even when pressure anisotropy is included, implying that anisotropy alone cannot explain the observed departures from ideal rotational discontinuities. Finally, the Alfvénicity of the current sheets correlates strongly with the cross helicity and residual energy of the surrounding solar wind, demonstrating that current sheet dynamics are largely governed by the ambient turbulent environment.
\end{abstract}

\begin{tcolorbox}[enhanced jigsaw, colback=white, opacitybacktitle=0.6, opacityback=0, leftrule=.75mm, toprule=.15mm, left=2mm, colframe=quarto-callout-note-color-frame, breakable, colbacktitle=quarto-callout-note-color!10!white, toptitle=1mm, titlerule=0mm, bottomtitle=1mm, title=\textcolor{quarto-callout-note-color}{\faInfo}\hspace{0.5em}{Note}, arc=.35mm, rightrule=.15mm, bottomrule=.15mm, coltitle=black]

This document and the accompanying website may contain outdated information.

\end{tcolorbox}

@@

Figure~\ref{fig-overview} provides an overview of the selected time interval from Encounter 7, comparing measurements from the PSP (left panels) and Wind (right panels) missions. From top to bottom, the panels display 8-minute averages of the magnetic field magnitude, proton number density, and proton bulk flow speed, followed by ion and electron temperatures resolved in the parallel and perpendicular directions. To facilitate comparison, Wind data are scaled to \(20 R_\odot\) using the empirical relations \(B_{\text{scaled}}/B = ( 20 R_\odot / 1\ \text{AU} )^{-1.59}\), \(n_{\text{scaled}}/n = ( 20 R_\odot / 1\ \text{AU} )^{-1.96}\) \citep{perroneRadialEvolutionSolar2019}.
The fifth and sixth panels show the median Alfvén ratio, \(R_{VB} = |Δ 𝐕| / |Δ 𝐕_A|\), of the current sheets and the total number of current sheets identified within each four-hour window. Including the anisotropy factor \(Λ = μ(p_{\parallel,e} − p_{\perp,e} + p_{\parallel,i} − p_{\perp,i})/B^2\) does not significantly affect the Alfvén ratio \(R_{VB}^* = R_{VB} / \sqrt{1 - Λ}\); its average value is approximately 0.5 for PSP and 0.4 for Wind over the entire time interval. The number of current sheets varies from 100 to 300 per four-hour window, corresponding to an average of about 1,000 and 900 current sheets per day for PSP and Wind, respectively. The PSP occurrence rates are broadly consistent with previous studies at similar heliocentric distances \citep{lotekarKineticscaleCurrentSheets2022}. We detect significantly more current sheets than the ∼150 events reported at 1 AU by \citet{vaskoKineticscaleCurrentSheets2022, zhangSolarWindDiscontinuities2025b}, largely because our detection is not restricted to only the most intense kinetic-scale current sheets (see methods below).
Panels (a.7) and (b.7) present the cross helicity and residual energy, calculated every 8 minutes, which characterize the Alfvénic nature and energetic balance of the ambient plasma. We compute these quantities in the standard form: \(\sigma_c = \frac{2 \langle \delta \mathbf{V} \cdot \delta \mathbf{V}_A \rangle}{ \langle\delta\mathbf{V}^2\rangle + \langle\delta\mathbf{V}_A^2\rangle }, \quad \sigma_r = \frac{\langle\delta \mathbf{V}^2\rangle-\langle\delta \mathbf{V}_A^2\rangle}{\langle\delta \mathbf{V}^2\rangle+\langle\delta \mathbf{V}_A^2\rangle},\) where \(\delta \mathbf{V}\) and \(\delta \mathbf{V}_A=\delta( \mathbf{B} / \sqrt{\mu_0 \rho})\) are the fluctuating velocity and Alfvén velocity, respectively.
The final two panels display 8-minute averages of helium abundance and plasma beta, providing complementary information on the compositional and thermal conditions of the solar wind. These parameters offer critical insights into identifying the source regions of different solar wind streams \citep{altermanCrossHelicityHelium2025}.
